\MakeUppercase{Graph Product Structure for Non-Minor-Closed Classes}

### 2022-03-07 k-planar.tex
Pfffft

  Begin with any width-$t$ tree-decomposition $(B_x:x\in V(T_0))$ of $H$ that uses some tree $T_0$.  Select any node $x\in V(T_0)$, add a leaf $x_0$, with $B_{x_0}=\emptyset$, adjacent to $x$ and root $T_0$ at $x_0$. (The purpose of $x_0$ is to ensure that every node $x$ for which $B_x$ is non-empty has a parent.)  Let $f:V(H)\to V(T)$ be the function that maps each $x\in V(H)$ onto the root of the subtree $T_0[x]:=T_0[\{y\in V(T_0): x\in B_y\}]$.  If $f$ is not one-to-one, then select some distinct pair $x,y\in V(H)$ with $a:=f(x)=f(y)$.  Subdivide the edge between $a$ and its parent in $T_0$ by introducing a new node $a'$ with $B_{a'}=B_{a}\setminus\{x\}$.  This modification reduces the number of distinct pairs $x,y\in V(H)$ with $f(x)=f(y)$, so repeatedly performing this modification will eventually produce a tree-decomposition $(B_x:x\in V(T_0))$ of $H$ in which $f$ is one-to-one.

@@

\note{DW}{Use $\SS$ for the given shortcut system and $\SS$ for the partition?}

@@

\referee{2}{Page 8.  Can you add a brief explanation why $\mathcal{P}$ is a
	partition of $V(G)$?  It is clear that the sets in $\mathcal{P}$ are
	disjoint, but why do they cover $V(G)$?}
\note{PM}{I added an explanation in a footnote since this is obvious.  In particular, for any function $f$, $\{f^{-1}(x):x\in\mathop{codomain}(f)\}$ is a partition of the $\mathop{domain}(f)$.}

From this point onward, the plan is to show that:
\begin{compactenum}[(i)]
  \item $\mathcal{P}$ has small layered width with respect to the layering $\mathcal{L}$ and that
  \item $J$ has small treewidth.
\end{compactenum}

\referee{1}{7. page 8, item (i): ``$S$ has small layered width with respect to the layering $L$'' $\longrightarrow$ ``$S$ has small layered width with respect to the layering $L$ of $G$''. Perhaps mention something about how it is easy to find the new layering.}

@@

\begin{proof}
  Recall that $S_x$ is defined by vertices that participate in $x$, and these are vertices that are either in $Y_x$ or in shortcuts that cross $Y_x$.  We say that a vertex $w\in Y_x$ \defin{contributes} a vertex $v\in S_x$ if $v$ participates in $x$.
  We upper bound the number of vertices in $S_x\cap L_i$ by upper-bounding the number of vertices contributed to $S_x\cap L_i$ by each $w\in Y_x$. Refer to \cref{contribute}.  If $w\in Y_x\cap L_i$ and no path in $\SS$ includes $w$ as an internal vertex then $w$ contributes at most one vertex, itself, to $S_x\cap L_i$.

\referee{1}{9. page 8, line 10: ``We say that a vertex $w \in Y_x$ contributes a vertex
  $v \in S_x$ if $v$ participates in $x$.'' This needs to be fixed.}

\note{DW}{What is wrong here?}

@@

  Otherwise, consider some path $P\in\SS$ that contains $w$ as an internal vertex.  If $w\in L_{i}$, then $P$ contributes at most $k+1$ vertices to $S_x\cap L_i$.  If $w\in L_{i-1}\cup L_{i+1}$, then $P$ contributes at most $k$ vertices to $S_x\cap L_i$. If $w\in L_{i-j}\cup L_{i+j}$ for $j\ge 2$, then $P$ contributes at most $k-j$ vertices to $S_x\cap L_i$.

>>>>

  Begin with any width-$t$ tree-decomposition $(B_x:x\in V(T_0))$ of $H$ that uses some tree $T_0$.  Select any node $x\in V(T_0)$, add a leaf $x_0$, with $B_{x_0}=\emptyset$, adjacent to $x$ and root $T_0$ at $x_0$. (The purpose of $x_0$ is to ensure that every node $x$ for which $B_x$ is non-empty has a parent.)  Let $f:V(H)\to V(T)$ be the function that maps each $x\in V(H)$ onto the root of the subtree $T_0[x]:=T_0[\{y\in V(T_0): x\in B_y\}]$.  If $f$ is not one-to-one, then select some distinct pair $x,y\in V(H)$ with $a:=f(x)=f(y)$.  Subdivide the edge between $a$ and its parent in $T_0$ by introducing a new node $a'$ with $B_{a'}=B_{a}\setminus\{x\}$. Now $f(y)=a'$ and $f(x)=a$, so this modification reduces the number of distinct pairs $x,y\in V(H)$ with $f(x)=f(y)$, so repeatedly performing this modification will eventually produce a tree-decomposition $(B_x:x\in V(T_0))$ of $H$ in which $f$ is one-to-one.

@@

\note{DW}{Use $\SS$ for the given shortcut system and $\SS$ for the partition?}
\note{PM}{Yes, I've done this.  Again, I was carefuly, but need to reread.}

@@

% \referee{2}{Page 8.  Can you add a brief explanation why $\mathcal{P}$ is a
% 	partition of $V(G)$?  It is clear that the sets in $\mathcal{P}$ are
% 	disjoint, but why do they cover $V(G)$?}
% \note{PM}{I added an explanation in a footnote since this is obvious.  In particular, for any function $f$, $\{f^{-1}(x):x\in\operatorname{range}(f)\}$ is a partition of $\operatorname{domain}(f)$.}

From this point onward, the plan is to show that:
\begin{compactenum}[(i)]
  \item $\mathcal{P}$ has small layered width with respect to the layering $\mathcal{L}$ of $G$ and that
  \item $J$ has small treewidth.
\end{compactenum}
Once we have established (i) and (ii), the result follows easily since a layering of $G^\SS$ is easily obtained from $\mathcal{L}$ by `compressing' groups of $k$ consecutive layers.
% \referee{1}{7. page 8, item (i): ``$P$ has small layered width with respect to the layering $L$'' $\longrightarrow$ ``$P$ has small layered width with respect to the layering $L$ of $G$''.}
% \referee{1}{Perhaps mention something about how it is easy to find the new layering.}

@@

\begin{proof}
  Recall that $S_x$ is defined by vertices that participate in $x$, and these are vertices that are either in $Y_x$ or in shortcuts that cross $Y_x$.  We say that a vertex $w\in Y_x$ \defin{contributes} a vertex $v\in S_x$ if $v=w$ or if some path in $\SS$ that contains $v$ has $w$ as an internal vertex.
  We upper bound the number of vertices in $S_x\cap L_i$ by upper-bounding the number of vertices contributed to $S_x\cap L_i$ by each $w\in Y_x$.

  Refer to \cref{contribute}.  If $w\in Y_x\cap L_i$ and no path in $\SS$ includes $w$ as an internal vertex then $w$ contributes at most one vertex, itself, to $S_x\cap L_i$.
  Otherwise, consider some path $P\in\SS$ that contains $w$ as an internal vertex.  If $w\in L_{i}$, then $P$ contributes at most $k+1$ vertices to $S_x\cap L_i$.  If $w\in L_{i-1}\cup L_{i+1}$, then $P$ contributes at most $k$ vertices to $S_x\cap L_i$. If $w\in L_{i-j}\cup L_{i+j}$ for $j\ge 2$, then $P$ contributes at most $k-j$ vertices to $S_x\cap L_i$.
% 
% \referee{1}{9. page 8, line 10: ``We say that a vertex $w \in Y_x$ contributes a vertex
%   $v \in S_x$ if $v$ participates in $x$.'' This needs to be fixed.}
% 
% \note{DW}{What is wrong here?}

@@
